### 2022-05-29 manuscript.Rmd
First Changes
Moved Algorithm part below. Added the starting sentence to that paragraph. Wrote a short paragraph on starting values.

	* Consider alternative parameterizations; e.g., free variances/covariances 
	* See Nylund
1. Starting values [@nylund-gibsonTenFrequentlyAsked2018; ]
    In many circumstances and especially in a cluster analysis setting, the mixture component means are expected to be different. Thus, a reasonable and largely employed way of initiating EM consists of starting from the solution of a K-means type algorithm. [@biernackiChoosingStartingValues2003]
1. Algorithm
    EM is a local optimizer. Imagine we are estimating two parameters, e.g. the class-specific means $\mu_c$ on a continuous indicator (ignoring the variance for now). The EM algorithm will attempt to find a combination of values for these two parameters that maximizes the likelihood ($LL$) of all observed data. In practice, instead of maximizing $LL$, often $-2*LL$ is minimized, as this offers computational advantages. We can think of this optimization problem as a three-dimensional landscape: The X and Y dimensions are determined by the class-specific means, so $X = \mu_1$ and $Y = \mu_2$ - and the Z-dimension is determined by $Z = -2*LL$. The optimizer must find the deepest "valley" in this landscape, which reflects the combination of $\mu_1$ and $\mu_2$ that maximizes the likelihood of the data.
The EM optimizer behaves somewhat like a marble, dropped in this landscape.
It is dropped at some random point in space, and will roll into the nearest valley.
The problem is that, once EM rolls into a valley, it will settle on the bottom of that valley (this is known as "convergence").
It cannot climb out again.
Thus, if their are multiple valleys, the risk is that the optimizer gets stuck in a shallower valley (a "local optimum"), and never discovers the deepest valley (the "global optimum", or best solution).
One solution to this problem is to drop many marbles at random places, compare their final $-2*LL$ values,
choose the solution with the lowest $-2*LL$, and make sure that several marbles replicated this solution.
This is the "random starts" approach.

@@

@@

>>>>

	* Consider alternative parameterizations; e.g., free variances/covariances 
	* See Nylund

@@

    Mixture model parameters are estimated through maximum likelihood (ML) estimation with expectation-maximization (EM) as a local optimizer. Imagine we are estimating two parameters, e.g. the class-specific means $\mu_c$ on a continuous indicator (ignoring the variance for now). The EM algorithm will attempt to find a combination of values for these two parameters that maximizes the likelihood ($LL$) of all observed data. In practice, instead of maximizing $LL$, often $-2*LL$ is minimized, as this offers computational advantages. We can think of this optimization problem as a three-dimensional landscape: The X and Y dimensions are determined by the class-specific means, so $X = \mu_1$ and $Y = \mu_2$ - and the Z-dimension is determined by $Z = -2*LL$. The optimizer must find the deepest "valley" in this landscape, which reflects the combination of $\mu_1$ and $\mu_2$ that maximizes the likelihood of the data.
The EM optimizer behaves somewhat like a marble, dropped in this landscape.
It is dropped at some random point in space, and will roll into the nearest valley.
The problem is that, once EM rolls into a valley, it will settle on the bottom of that valley (this is known as "convergence").
It cannot climb out again.
Thus, if their are multiple valleys, the risk is that the optimizer gets stuck in a shallower valley (a "local optimum"), and never discovers the deepest valley (the "global optimum", or best solution).
One solution to this problem is to drop many marbles at random places, compare their final $-2*LL$ values,
choose the solution with the lowest $-2*LL$, and make sure that several marbles replicated this solution.
This is the "random starts" approach.

@@

The algorithm needs starting values for each parameter it wants to estimate. If it were provided a single starting value for each parameter, the  EM optimizer would be at risk of finding a local optimum instead of the global optimum. In other words, it would converge on a suboptimal solution and the resulting parameter estimates would be biased. To prevent this, we should give the algorithm a sufficient set of random starting values for each parameter in order to maximize our chances of finding the deepest valley that is the global optimum (Geiser, 2013).

[@nylund-gibsonTenFrequentlyAsked2018; ]

    In many circumstances and especially in a cluster analysis setting, the mixture component means are expected to be different. Thus, a reasonable and largely employed way of initiating EM consists of starting from the solution of a K-means type algorithm. [@biernackiChoosingStartingValues2003]
